What's in a Name? Linear Temporal Logic Literally Represents Time Lines

### 2023-01-01 paper/main.tex
small update to main
added metric defs

@@

    \begin{highlights} %TODO ~ Highly encouraged but not mandatory
        \item Research highlight 1
        \item Research highlight 2
    \end{highlights}

@@

\noindent
\begin{tabular}{p{3.6in}|c|c}
  \hline
  Purpose & \# formulas & Source \\
  \hline
  Design space for NASA's AAC (Automated Airspace Concept) & 20,250 & \cite{GCMTR16} \\
  \hline
  Safety specificatons for Boeing's AIR6110 wheel brake system & & \cite{BCFJKPRT15} \\
  \hline
  Requirements for model checking NASA's AAC high-level architecture & 6 & \cite{ZR14} \\
  \hline
  Real formulas collected from all publicly-available, pre-2011 case studies &  & \cite{RV11} \\
  \hline
\end{tabular}

@@

KYR: Define size, length, height, complexity, etc. here.

>>>>

\newcommand{\tllen}{\text{tllen}}

@@

    \begin{highlights} 
        \item We transform Linear Temporal Logic (LTL) formulas to timeline visualizations for the set of satisfying traces.
        \item We translate LTL formulas to \Buchi automata, then to $\omega$-regular expressions, then simplify them and use graph visualization tools to draw timelines.
        \item Our algorithm is sound and complete. 
        \item Our tool, \texttt{ltl2timeline}, effectively constructs small timeline visualizations for a range of LTL formulas.
    \end{highlights}

@@

\noindent
\begin{tabular}{p{3.6in}|c|c}
  \hline
  Purpose & \# formulas & Source \\
  \hline
  Design space for NASA's AAC (Automated Airspace Concept) & 20,250 & \cite{GCMTR16} \\
  \hline
  Safety specificatons for Boeing's AIR6110 wheel brake system & & \cite{BCFJKPRT15} \\
  \hline
  Requirements for model checking NASA's AAC high-level architecture & 6 & \cite{ZR14} \\
  \hline
  Real formulas collected from all publicly-available, pre-2011 case studies & 63 & \cite{RV11} \\
  \hline
\end{tabular}

@@

%KYR: Define size, length, height, complexity, etc. here.

We use two measures of complexity for timeline visualizations: timeline length and star height.

We use timeline length, defined below, to capture the length of our timeline visualization graphs. Since Kleene stars are the trickiest operator to depict in timelines, we also use star height, a well-studied measure for the structural complexity of regular expressions, to measure the complexity of our visualizations.

\begin{definition}[Timeline Length]
    We define the timeline length of a regular expression $\tllen(A)$ recursively as follows
    \begin{align*}
        \tllen(\emptyset)   &= -\infty\\
        \tllen(\epsilon)        &= 0\\
        \tllen(a)           &= 1\\
        \tllen(A_1 A_2)     &= \tllen(A_1) + \tllen(A_2)\\
        \tllen(A_1 + A_2)   &= \max(\tllen(A_1, A_2))\\
        \tllen(A^*)         &= 2\tllen(A) + 1 
    \end{align*}
    We also extend this definition to $\omega$-regular expressions as follows
    \begin{align*}
        \tllen(A^\omega)    &= \tllen(A)\\
        \tllen(AB)          &= \tllen(A) + \tllen(B)\\
        \tllen(B_1 + B_2)   &= \max(\tllen(B_1), \tllen(B_2))
    \end{align*}
    This quantity $\tllen$ captures the length of the longest path in our timeline graph visualization.
\end{definition}

\begin{definition}[Star height]
    We define the star height of regular and $\omega$-regular expressions to be the star-nesting depth in the (unsimplified) expression. Namely,
    \begin{align*}
        h(\emptyset) = h(\epsilon) = h(a) &= 0\\
        h(A_1 A_2) = h(A_1 + A_2) &= \max(h(A_1), h(A_2))\\
        h(A^*) &= 1 + h(A)\\
        h(A^\omega) &= h(A)\\
        h(A B) &= \max(h(A), h(B))\\
        h(B_1 + B_2) &= \max(h(B_1), h(B_2))
    \end{align*}

\end{definition}
